On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-08-25 Manuscript/manuscript.tex
removed addressed comments and one new comment

\paragraph{Percentile CI}
The percentile CI was first suggested in \citet{efron1979bootstrap}.
\jy{The previous superscript does not make sense because
  $b \in \{1, \ldots,  B\}$. Make similar changes to other places as needed.} 
\mc{addressed}
Let $\hat\theta_{n, \alpha}^B$ be the empirical $100\alpha$th percentile of
$\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$. 
A $(1 - \alpha)100\%$ empirical percentile CI is
\[
(\hat\theta_{n, \alpha/2}^{B}, \quad \hat\theta_{n, 1 - \alpha/2}^{B}).
\]
This CI is not necessarily centered by the original point estimate 
$\hat\theta_n$. As will be shown in our simulation study, this approach works 
well for the marginal mean and standard deviation of a serially dependent 
process, but its coverage of the temporal dependence deteriorates as $n$ 
increases, which is contrary to what one would expect.

\paragraph{Bias-Corrected (BC) CI}
The procedure for constructing a bias-corrected Bootstrap CI based on standard
bootstrap is described in \citet{carpenter2000bootstrap}. Let
$b = \Phi^{-1}\{\#\{\hat\theta_n^{(b)} < \hat{\theta}_n\} / B\}$ for
$b \in \{1, \ldots, B\}$. Define $\alpha_1 = \Phi(2b - z_{\alpha/2})$ and 
$\alpha_2 = \Phi(2b - z_{1 - \alpha/2})$. 
A $(1 - \alpha)100\%$ ~BC CI is
\[
(\hat\theta_{n, \alpha_1}^{B}, \quad \hat\theta_{n, \alpha_2}^{B}).
\]

@@

We propose a CI that is centered at the original point estimate and uses the 
variation in the bootstrap estimates to construct the error bound. This interval 
requires
\jy{Note the notation change and do a thorough check} \mc{addressed}
the computation of $\bar\theta_n^B = n^{-1}\sum_{b=1}^B \hat\theta_n^{(b)}$, the 
mean of all bootstrap point estimates, and 
$\widehat{\text{bias}} = \bar\theta_n^{B} - \hat\theta_n$. 
A $(1 - \alpha)100\%$ CI is
\[
(\hat\theta_{n, \alpha/2}^{B} - \widehat{\text{bias}}, \quad
\hat\theta_{n, 1 - \alpha/2}^{B} - \widehat{\text{bias}}).
\]
This CI is centered around $\hat\theta_n$ and can also be rewritten as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{B} - \bar\theta_n^{B}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{B} - \bar\theta_n^{B}).
\]
It is not necessarily symmetric, as different critical values are used to 
compute the lower and upper bounds. It has the same width as the percentile CI.

@@

For estimating the standard deviation parameter~$\sigma_x$,
Figure~\ref{fig:sigma} suggests that every method can reach nominal coverage of
$\sigma_x$ if the sample is large enough, but for a given $n$ and $\phi$,
coverage of $\sigma_x$ will be lower than coverage of $\mu$ in general. Like 
$\mu$, $\sigma_x$ can be covered by Student $t$ CIs with smaller sample sizes
when compared to other methods. Unlike $\mu$, there is no over-coverage issue 
for $\sigma_x$ when $\phi = -0.4$. Standard, percentile, BC, BCA, and recentered 
percentile CIs again have similar performance. All methods seem to have slightly 
higher coverage of $\sigma_x$ when $\phi$ is negative versus when $\phi$ is 
positive. Regardless of the sign, coverage of $\sigma_x$ gets worse as the 
strength of the temporal dependence increases.
\jy{Compared to Figure 1, there is no over-coverage in this case. Say something
  about this observation.} \mc{addressed}

@@

\jy{check for no uppercases for recentered percentile, etc. throughout}
\mc{addressed}

@@

\jy{The red cross is not $\theta_n^{(b)}$, but their average over
  $b = 1,  \ldots, B$. Define this as bootstrap point estimate in section 2.2
  for ease of referencing in the discussion.}
\mc{I'm not sure if I addressed this like you intended, I defined 
$\hat\theta_n^{(b)}$ as bootstrap point estimate, and referred to 
$\hat\theta_n$ as the original point estimate. Did you want me to define
$\bar\theta_n^{B}$ as the bootstrap point estimate?}

\jy{Keep the interval labels consistent in figure and in the text}
\mc{is it fine to have them capitalized in the figures?}

The outcomes of the $\phi$ estimation raise a natural question about the 
lackluster performance of certain methodologies. To delve into this inquiry, a
set of 50 CIs was generated for each of the percentile, BC, and recentered
percentile  approaches for samples of $n \in \{800, 1600\}$. Illustrated in 
Figure~\ref{fig:npi}, it becomes evident that the percentile-based CIs exhibit a 
notable bias, predominantly manifesting as a substantial underestimation of 
$\phi$ with point estimator $\bar\theta_n^B$, that is, the average of $B$ 
bootstrap point estimates. As the sample size increases from 800 to 1600, this 
bias does not vanish while the uncertainty reduces, which explains why the 
coverage rates deteriorate. The bias in $\bar\theta_n^B$ also appears to 
invalidate the bias-correction in the BC bootstrap, leading to the poor 
performance of the BC intervals. The BCA intervals have the same problem as the 
BC intervals in the bias-correction step. The root of the issue appears to be 
that the autocorrelation in the block bootstrap samples is somehow smaller 
compared to that in the original sample. On the other hand, the original point 
estimate $\hat\theta_n$ is asymptotically unbiased. Since the width is based on 
the uncertainty in the bootstrap point estimates $\hat\theta_n^{(b)}$, 
$b = 1, \ldots, B$, the percentile CIs recentered at the original point estimate 
provide desired coverage.

@@

\jy{use a better topic sentence.} \mc{addressed}
One of the goals of this study was to provide some practical recommendations for 
necessary sample sizes when using block bootstrap to estimate parameters of 
serially dependent data. When using Student's $t$ intervals and $\phi$ is 
unknown, the results of this study suggest that an $n \geq 800$ may be necessary 
for common practice to estimate simpler parameters like $\mu$ or $\sigma_x$. If 
$\phi$ is already known, a lesser $n$ may be adequate to estimate these 
parameters. For lower absolute values of $\phi$, a smaller $n$ is sufficient 
when estimating $\sigma_x$. For negative values of $\phi$, coverage of all 
parameters is higher in general. In fact, Student's $t$ seems to over-cover 
$\mu$ for $\phi = - 0.4$ when using smaller sample sizes, meaning that if $\phi$ 
is known to have a very negative value, another CI may be required. However, in 
real world applications, $\phi$ is more commonly found to be positive, so 
Student's $t$ should still be used if $\phi$ is unknown. Lastly, to estimate 
$\phi$, $n \geq 100$ using the Student's $t$ method may be sufficient. Further
investigation may be necessary to see if there are other interval corrections
that fix the coverage deterioration problem for $\phi$.

This study could be used as a guide for applied statistics courses for students
to generally understand how large of a sample size is sufficient for block
bootstrap to be used versus other inference methods. For undergraduate or 
graduate students, block bootstrap is not typically a part of curriculum, but
the results of this study can easily be used to demonstrate when it is practical 
to use this method. This information could also prove to be useful for research 
using block bootstrap estimation of time series in domains such as econometrics. 
Future studies could investigate the $n$ needed to make inferences about other 
forms of serially dependent data such as a moving average process. One could 
also investigate if there are types of block bootstrap interval construction 
such as $ABC$ or bootstrap-t intervals, or alternatives to block bootstrap, such 
as AR-Sieve bootstrap \citep{kreiss1992bootstrap}, that could more
\jy{Cite references for AR-sieve} \mc{addressed}
appropriately recover the parameters of a time series. In addition, as noted by
\citet{buhlmann2002bootstraps}, there are some general drawbacks of block 
bootstrap - with respect to how reasonably it imitates the data-generating 
process - which may be addressed in a future study.

>>>>

\paragraph{Percentile CI}
The percentile CI was first suggested in \citet{efron1979bootstrap}.
Let $\hat\theta_{n, \alpha}^B$ be the empirical $100\alpha$th percentile of
$\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$. 
A $(1 - \alpha)100\%$ empirical percentile CI is
\[
(\hat\theta_{n, \alpha/2}^{B}, \quad \hat\theta_{n, 1 - \alpha/2}^{B}).
\]
This CI is not necessarily centered by the original point estimate 
$\hat\theta_n$. As will be shown in our simulation study, this approach works 
well for the marginal mean and standard deviation of a serially dependent 
process, but its coverage of the temporal dependence deteriorates as $n$ 
increases, which is contrary to what one would expect.

\paragraph{Bias-Corrected (BC) CI}
The procedure for constructing a bias-corrected Bootstrap CI based on standard
bootstrap is described in \citet{carpenter2000bootstrap}. Let
\jy{Use a different symbol because $b$ has been used. Isn't is the same as $\hat
  z_0$ in the next paragraph?}
$b = \Phi^{-1}\{\#\{\hat\theta_n^{(b)} < \hat{\theta}_n\} / B\}$ for
$b \in \{1, \ldots, B\}$. Define $\alpha_1 = \Phi(2b - z_{\alpha/2})$ and 
$\alpha_2 = \Phi(2b - z_{1 - \alpha/2})$. 
A $(1 - \alpha)100\%$ ~BC CI is
\[
(\hat\theta_{n, \alpha_1}^{B}, \quad \hat\theta_{n, \alpha_2}^{B}).
\]

@@

We propose a CI that is centered at the original point estimate and uses the 
variation in the bootstrap estimates to construct the error bound. This interval 
requires
the computation of $\bar\theta_n^B = n^{-1}\sum_{b=1}^B \hat\theta_n^{(b)}$, the 
mean of all bootstrap point estimates, and 
$\widehat{\text{bias}} = \bar\theta_n^{B} - \hat\theta_n$. 
A $(1 - \alpha)100\%$ CI is
centered around $\hat\theta_n$ and can be rewritten as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{B} - \bar\theta_n^{B}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{B} - \bar\theta_n^{B}).
\]
It is not necessarily symmetric, as different critical values are used to 
compute the lower and upper bounds. It has the same width as the percentile CI.

@@

For estimating the standard deviation parameter~$\sigma_x$,
Figure~\ref{fig:sigma} suggests that every method can reach nominal coverage of
$\sigma_x$ if the sample is large enough, but for a given $n$ and $\phi$,
coverage of $\sigma_x$ will be lower than coverage of $\mu$ in general. Like 
$\mu$, $\sigma_x$ can be covered by Student $t$ CIs with smaller sample sizes
when compared to other methods. Unlike $\mu$, there is no over-coverage issue 
for $\sigma_x$ when $\phi = -0.4$. Standard, percentile, BC, BCA, and recentered 
percentile CIs again have similar performance. All methods seem to have slightly 
higher coverage of $\sigma_x$ when $\phi$ is negative versus when $\phi$ is 
positive. Regardless of the sign, coverage of $\sigma_x$ gets worse as the 
strength of the temporal dependence increases.

@@

@@

The outcomes of the $\phi$ estimation raise a natural question about the 
lackluster performance of certain methodologies. To delve into this inquiry, a
set of 50 CIs was generated for each of the percentile, BC, and recentered
percentile  approaches for samples of $n \in \{800, 1600\}$. Illustrated in 
Figure~\ref{fig:npi}, it becomes evident that the percentile-based CIs exhibit a 
notable bias, predominantly manifesting as a substantial underestimation of 
$\phi$ with point estimator $\bar\theta_n^B$, that is, the average of $B$ 
bootstrap point estimates. As the sample size increases from 800 to 1600, this 
bias does not vanish while the uncertainty reduces, which explains why the 
coverage rates deteriorate. The bias in $\bar\theta_n^B$ also appears to 
invalidate the bias-correction in the BC bootstrap, leading to the poor 
performance of the BC intervals. The BCA intervals have the same problem as the 
BC intervals in the bias-correction step. The root of the issue appears to be 
that the autocorrelation in the block bootstrap samples is somehow smaller 
compared to that in the original sample. On the other hand, the original point 
estimate $\hat\theta_n$ is asymptotically unbiased. Since the width is based on 
the uncertainty in the bootstrap point estimates $\hat\theta_n^{(b)}$, 
$b = 1, \ldots, B$, the percentile CIs recentered at the original point estimate
$\hat\theta_n$ provide desired coverage.

@@

One of the goals of this study was to provide some practical recommendations for 
necessary sample sizes when using block bootstrap to estimate parameters of 
serially dependent data. When using Student's $t$ intervals and $\phi$ is 
unknown, the results of this study suggest that an $n \geq 800$ may be necessary 
for common practice to estimate simpler parameters like $\mu$ or $\sigma_x$. If 
$\phi$ is already known, a lesser $n$ may be adequate to estimate these 
parameters. For lower absolute values of $\phi$, a smaller $n$ is sufficient 
when estimating $\sigma_x$. For negative values of $\phi$, coverage of all 
parameters is higher in general. In fact, Student's $t$ seems to over-cover 
$\mu$ for $\phi = - 0.4$ when using smaller sample sizes, meaning that if $\phi$ 
is known to have a very negative value, another CI may be required. However, in 
real world applications, $\phi$ is more commonly found to be positive, so 
Student's $t$ should still be used if $\phi$ is unknown. Lastly, to estimate 
$\phi$, $n \geq 100$ using the Student's $t$ method may be sufficient. Further
investigation may be necessary to see if there are other interval corrections
that fix the coverage deterioration problem for $\phi$.

This study could be used as a guide for applied statistics courses for students
to generally understand how large of a sample size is sufficient for block
bootstrap to be used versus other inference methods. For undergraduate or 
graduate students, block bootstrap is not typically a part of curriculum, but
the results of this study can easily be used to demonstrate when it is practical 
to use this method. This information could also prove to be useful for research 
using block bootstrap estimation of time series in domains such as econometrics. 
Future studies could investigate the $n$ needed to make inferences about other 
forms of serially dependent data such as a moving average process. One could 
also investigate if there are types of block bootstrap interval construction 
such as $ABC$ or bootstrap-t intervals, or alternatives to block bootstrap, such 
as AR-Sieve bootstrap \citep{kreiss1992bootstrap}, that could more
appropriately recover the parameters of a time series. In addition, as noted by
\citet{buhlmann2002bootstraps}, there are some general drawbacks of block 
bootstrap --- with respect to how reasonably it imitates the data-generating 
process --- which may be investigated in a future study.
